---
title: "simpleRegress"
author: "Aetan Barquero"
date: "Fall 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
simp <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

The predicted weight of a bear compared to a chest has an adjusted R-Squared value of 0.9262, and residual standard error value of 33.08.

```{r}
simp <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(simp)
summary.lm(simp)
```

The predicted overall weight of a bear compared to length has a residual standard errror of 61.83 and an r-squared value of 0.7472.

```{r}
simp <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```
The predicted weight of a bear compared to head width has a r squared value of 0.6138 and a residual standar error value of 76.41

```{r}
simp <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```

The predicted weight of a bea compared to head length has a r squared value of 0.6959 and a residual standard error of 67.81

```{r}
simp <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(simp)
summary.lm(simp)
```

The predicted weight of a bear compared to age has a r squared value of 0.561 and a residual standard error value of 81.47

# CONCLUSION AND RANKING



With the collected data, I am able to rank the compared weights:

1. Weight~CHEST RSE: 33.08 and ARS: 0.9262

2. Weight~NECK RSE: 43.91 and ARS: 0.87

3. Weight~Length RSE: 61.83 and ARS: 0.7423

4. Weight~Headlen RSE: 67.81 and ARS: 0.69

5. Weight~Headwth RSE: 76.41 and ARS: 0.6064

6. Weight~AGE RSE: 81.47 and ARS: 0.5526